\bf On semidefinite representable sets over valuation fields

### 2024-11-13 article/paper_padic_spec.tex
commit by Simone

\begin{abstract}
  \noindent
  Real polyhedra and real spectrahedra (or their projections) are respectively the feasible sets
  of linear programming (LP) and semidefinite programming (SDP). 
  This paper investigates versions of these sets and of
  the related optimization problems with data over a field $K$ with valuation.
  We showcase examples of non-polyhedral $K$-spectrahedra and of sets that are semidefinite
  representable over $K$ but are not $K$-spectrahedra.
\end{abstract}

@@

\begin{remark}
  As in the real case, one often abuses of notation, saying that the preimage
  $S = \varphi^{-1}(K^{d \times d}_+) = \{x \in K^n : A_0+x_1A_1+\cdots+x_nA_n \succeq 0\}$
  of $K^{d \times d}_+$ under the map $\varphi(x) = A_0+x_1A_1+\cdots+x_nA_n$, is a \emph{spectrahedron
  in $K^n$}.
\end{remark}

\begin{remark}
  The class of spectrahedra is closed under intersection. Indeed, for affine spaces $\mathcal{L}_1,\mathcal{L}_2$
  one has $(\mathcal{L}_1 \cap K^{d \times d}_+) \cap (\mathcal{L}_2 \cap K^{d \times d}_+) =
  (\mathcal{L}_1 \cap \mathcal{L}_2) \cap K^{d \times d}_+$. Now
  assuming $\mathcal{L}_1 = A_0+\span{A_1,\ldots,A_n}$ and $\mathcal{L}_2 = B_0+\span{B_1,\ldots,B_n}$,
  then $S_A \cap S_B = S_{A \oplus B}$ (indeed both $A \succeq 0$ and $B \succeq 0$ if and only if
  $A \oplus B \succeq 0$).
\end{remark}

%\begin{remark}
%	As in the real case, the spectrahedron associated with the affine plan $\mathcal{L}$ will often be identified with its preimage in $K^s$ (i.e. the set of vectors $x \in K ^s$ so that the linear matrix $A(x) = A_0 + x_1A_1 + \ldots + x_sA_s$ is positive semidefinite with $A_0,A_1,\ldots,A_s$ spanning $\mathcal{L}$). 
%\end{remark}
\begin{proposition}
  A polyhedron is a spectrahedron.
\end{proposition}
\begin{proof}
  Let $\ell_i,m_j \in K[x]_1$, for $i=1,\ldots,d$ and $j=1,\ldots,e$, and let $P = Q \cap L \subset K^n$
  be a polyhedron as in \Cref{def_polyhedra}, for $Q = \{x \in K^n : \val(\ell_i) \geq 0, i=1,\ldots,d\}$
  and $L = \{x \in K^n : \val(m_j) = +\infty, j=1,\ldots,e\}$.
  First, denote by $A(x) = \ell_1 \oplus \cdots \oplus \ell_d \in K[x]^{d\times d}_1$,
  so that $A(x) = A_0+x_1A_1+\cdots+x_nA_n$ for some (diagonal) $A_i \in K^{d\times d}$.
  We might assume that $A_0,A_1,\ldots,A_n$ are affinely independent, so that the mapping
  $\varphi(x) = A(x)$ is injective. 
  First, remark that $Q=S_A$ and thus $\varphi(Q) = \mathcal{L} \cap K^{d\times d}$, with
  $\mathcal{L} = A_0+\span{A_1,\ldots,A_n}$ is a spectrahedron.
  Next, remark that $L \subset K^n$ is the affine space defined by equations $m_1=\ldots=m_e=0$,
  hence $\varphi(L) \subset K^{d\times d}$ is an affine space. Since $\varphi$ is injective,
  one has $\varphi(P) = \varphi(Q \cap L) = \varphi(Q) \cap \varphi(L)$ which is the intersection
  of a spectrahedron with an affine space, thus a spectrahedron.
  %Define $B = m_1 \oplus \cdots \oplus m_e \in K[x]^{e\times e}_1$
\end{proof}

>>>>

\begin{abstract}
  \noindent
  Real polyhedra and real spectrahedra (or their projections) are respectively the feasible sets
  of linear programming (LP) and semidefinite programming (SDP). 
  This paper investigates versions of these sets and of
  the related optimization problems with data over a field $K$ with valuation.
  DIRE CE QU'ON FAIT.
  We showcase examples of non-polyhedral $K$-spectrahedra and of sets that are semidefinite
  representable over $K$ but are not $K$-spectrahedra.
\end{abstract}

@@

\begin{remark}
  As in the real case, one often abuses of notation, saying that the preimage
  $S_A = \varphi_A^{-1}(K^{d \times d}_+) = \{x \in K^n : A_0+x_1A_1+\cdots+x_nA_n \succeq 0\}$
  of $K^{d \times d}_+$ under the map $\varphi_A : K^n \to K^{d \times d}$, defined as
  $\varphi_A(x) = A_0+x_1A_1+\cdots+x_nA_n$, is a \emph{spectrahedron
  in $K^n$}. Remark that the map $\varphi_A$ is in general not injective, unless $A_1,\ldots,A_n$ are linearly
  indepedent.
\end{remark}

\begin{remark}
  The class of spectrahedra in $K^{d \times d}$ (or in $K^n$) is closed under intersection: indeed one
  has $(\mathcal{L}_1 \cap K^{d \times d}_+) \cap (\mathcal{L}_2 \cap K^{d \times d}_+) =
  (\mathcal{L}_1 \cap \mathcal{L}_2) \cap K^{d \times d}_+$ and
  $S_A \cap S_B = S_{\diag(A,B)}$, for two affine spaces $\mathcal{L}_1 = A_0+\span{A_1,\ldots,A_n}$
  et $\mathcal{L}_2 = A_0+\span{B_1,\ldots,B_n}$.
\end{remark}

%\begin{remark}
%	As in the real case, the spectrahedron associated with the affine plan $\mathcal{L}$ will often be identified with its preimage in $K^s$ (i.e. the set of vectors $x \in K ^s$ so that the linear matrix $A(x) = A_0 + x_1A_1 + \ldots + x_sA_s$ is positive semidefinite with $A_0,A_1,\ldots,A_s$ spanning $\mathcal{L}$). 
%\end{remark}

\begin{proposition}
  A polyhedron in $K^{d \times d}$ is a spectrahedron in $K^{d \times d}$.
\end{proposition}
\begin{proof}
  Let $\ell_i,m_j \in K[x]_1$, for $i=1,\ldots,d$ and $j=1,\ldots,e$, and let $P = Q \cap L \subset K^n$
  be a polyhedron as in \Cref{def_polyhedra}, for $Q = \{x \in K^n : \val(\ell_i) \geq 0, i=1,\ldots,d\}$
  and $L = \{x \in K^n : \val(m_j) = +\infty, j=1,\ldots,e\}$.
  First, denote by $A(x) = \ell_1 \oplus \cdots \oplus \ell_d \in K[x]^{d\times d}_1$,
  so that $A(x) = A_0+x_1A_1+\cdots+x_nA_n$ for some (diagonal) $A_i \in K^{d\times d}$.
  We might assume that $A_0,A_1,\ldots,A_n$ are affinely independent, so that the mapping
  $\varphi(x) = A(x)$ is injective. 
  First, remark that $Q=S_A$ and thus $\varphi(Q) = \mathcal{L} \cap K^{d\times d}$, with
  $\mathcal{L} = A_0+\span{A_1,\ldots,A_n}$ is a spectrahedron.
  Next, remark that $L \subset K^n$ is the affine space defined by equations $m_1=\ldots=m_e=0$,
  hence $\varphi(L) \subset K^{d\times d}$ is an affine space. Since $\varphi$ is injective,
  one has $\varphi(P) = \varphi(Q \cap L) = \varphi(Q) \cap \varphi(L)$ which is the intersection
  of a spectrahedron with an affine space, thus a spectrahedron.
  %Define $B = m_1 \oplus \cdots \oplus m_e \in K[x]^{e\times e}_1$
\end{proof}
